ediblecity: an R package to model and estimate the benefits of urban agriculture

### 2022-11-18 paper_v01.Rmd
finish first version

@@

`r ref("Table", "structure")`: Structure of the urban representation example"

```{r}
tribble(
  ~Column, ~Description,
  "Function", "A category representing the urban elements",
  "Function_verbose", "A more detailed category for the element, for example, if it is residential",
  "floors", "The number of floors of the element, 0 for non-built",
  "area", "The surface of the element",
  "flat_area", "The surface of the element that is flat (slope < 5º)",
  "edible_area", "The surface that is used to grow edible plants. Only applicable to urban agriculture solutions"
) |> 
  kable()

Along with the example for a urban representation, the `ediblecity` package also includes a `data.frame` with the general attributes of each green typology in the urban representation used to estimate the indicators (`r ref("Table", "city_functions")`). However, the user can provide their own attributes to estimate any indicator. In `city_functions` there are other columns not shown in table `r cross_refs[["city_functions"]]`, they are logical variables (i.e. TRUE/FALSE) used internally by the package to select urban agriculture elements.

@@

`r cross_refs[["city_functions"]]`: General attributes of the elements of urban green used to estimate the indicators.

@@

city_functions |>
  select(-starts_with("...")) |> 
  select(-edible, -public, -jobs, -volunteers, -location) |>
  select(-starts_with("water")) |> 
  kable()
```
### Indicators estimated

@@

The `ediblecity` package also provides the user with a function to create new scenarios based on the urban representation and a predefined set of urban agriculture solutions (`r ref("Table", "scenarios")`) based on where they are located (private gardens, plots on ground or rooftops) and their purpose (private, community or commercial). The function returns a spatial representation of the new scenario (`sf` object) with the same structure of the urban representation.

`r cross_refs[["scenarios"]]`: Elements created in new scenarios.

@@

tibble(
  `Urban agriculture solutions` = city_functions$functions[1:5],
  Location = c("Private gardens", rep("Plots on ground", 2), rep("Rooftops", 2)),
  Purpose = c("Private", rep("Community", 2), rep("Commercial", 2))
) |> 
  kable()

@@

`r ref("Table", "count_scenarios")` shows the number of agriculture solutions in each scenario.

@@

counting <- function(x, .name){
  x |> 
    st_drop_geometry() |> 
    filter(Function %in% city_functions$functions[1:5]) |> 
    count(Function, name = .name)
}

counting(scenarios[[1]], "s0") |> 
  full_join(counting(scenarios[[2]], "s1")) |> 
  full_join(counting(scenarios[[3]], "s2")) |> 
  mutate(across(where(is.numeric), ~if_else(is.na(.x), 0L, .x))) |> 
  kable()
```

## Calculate indicators to compare scenarios

@@

# We use the SVF object that is provided for the city_example in the package
map_dfr(scenarios, UHI, SVF = ediblecity::SVF, .id = "Scenario") |> 
  kable()

@@

# plot <- map(scenarios, ~ UHI(.x, SVF = ediblecity::SVF, return_raster = TRUE), .id = "Scenarios") |> 
#   map(~ ggplot()+
#         stars::geom_stars(data = .x)+
#         coord_equal()+
#         scale_fill_distiller(palette = "RdYlBu")+
#         theme_void()+
#         theme(plot.title = element_text(hjust=0.5))
#       ) |> 
#   map2(names(scenarios), ~ .x+labs(title = .y)) |> 
#   wrap_plots()+plot_layout(guides = 'auto', nrow = 1)
# 
# ggsave("assets/UHI_plots.png")

@@

map_dfr(scenarios, runoff_prev, .id="scenario") |> 
  kable()

@@

@@

map_dfr(scenarios, no2_seq, .id="Scenario") |> 
  kable()
```

@@

scenarios |> 
  map_dfr(food_production, .id = "Scenario") |> 
  mutate(across(where(is.numeric), ~ .x/1000)) |> 
  kable()

@@

## Acknowledgements

`r stop("add acknowledgements to R commnunity")`

>>>>

```
\begin{center}

\end{center}

tribble(
  ~Column, ~Description,
  "Function", "A category representing the urban elements",
  "Function_verbose", "A more detailed category for the element, for example, if it is residential",
  "floors", "The number of floors of the element, 0 for non-built",
  "area", "The surface of the element",
  "flat_area", "The surface of the element that is flat (slope < 5º)",
  "edible_area", "The surface that is used to grow edible plants. Only applicable to urban agriculture solutions"
) |> 
  kable(caption = "Structure of the urban representation example")

@@

Along with the example for a urban representation, the `ediblecity` package also includes a `data.frame` with the general attributes of each green typology in the urban representation used to estimate the indicators (Table 2). However, the user can provide their own attributes to estimate any indicator. In `city_functions` there are other columns not shown in table `r cross_refs[["city_functions"]]`, they are logical variables (i.e. TRUE/FALSE) used internally by the package to select urban agriculture elements.

@@

@@

city_functions |>
  select(-starts_with("...")) |> 
  select(-edible, -public, -jobs, -volunteers, -location) |>
  select(-starts_with("water")) |> 
  kable(caption = "General attributes of the elements of urban green used to estimate the indicators")
```
### Indicators estimated

@@

The `ediblecity` package also provides the user with a function to create new scenarios based on the urban representation and a predefined set of urban agriculture solutions (Table 3) based on where they are located (private gardens, plots on ground or rooftops) and their purpose (private, community or commercial). The function returns a spatial representation of the new scenario (`sf` object) with the same structure of the urban representation.

@@

tibble(
  `Urban agriculture solutions` = city_functions$functions[1:5],
  Location = c("Private gardens", rep("Plots on ground", 2), rep("Rooftops", 2)),
  Purpose = c("Private", rep("Community", 2), rep("Commercial", 2))
) |> 
  kable(caption = "Elements created in new scenarios")

@@

@@

counting <- function(x, .name){
  x |> 
    st_drop_geometry() |> 
    filter(Function %in% city_functions$functions[1:5]) |> 
    group_by(Function) |> 
    summarize(!!sym(paste(.name, "n")) := n(),
              !!sym(paste(.name, "area")) := sum(area))
}

counting(scenarios[[1]], "s0") |> 
  full_join(counting(scenarios[[2]], "s1")) |> 
  full_join(counting(scenarios[[3]], "s2")) |> 
  mutate(across(where(is.numeric), ~if_else(is.na(.x), 0L, .x))) |> 
  relocate(Function, ends_with("n"), ends_with("area")) |> 
  kable(caption = "Number of agriculture solutions in each scenario and their surfaces")
```

## Calculate indicators and compare scenarios

@@

# We use the SVF object that is provided for the city_example in the package
map_dfr(scenarios, UHI, SVF = ediblecity::SVF, .id = "Scenario") |> 
  kable(caption = "Summary statistic of urban heat island effect in each scenario")

@@

# map(scenarios, ~ UHI(.x, SVF = ediblecity::SVF, return_raster = TRUE), .id = "Scenarios") |>
#   map(~ ggplot()+
#         stars::geom_stars(data = .x, na.action = na.omit)+
#         coord_equal()+
#         scale_fill_distiller(palette = "RdYlBu")+
#         theme_void()+
#         theme(plot.title = element_text(hjust=0.5))
#       ) |>
#   map2(names(scenarios), ~ .x+labs(title = .y)) |>
#   wrap_plots()+plot_layout(guides = 'auto', nrow = 1)
# 
# ggsave("assets/UHI_plots.png")

@@

map_dfr(scenarios, runoff_prev, .id="scenario") |> 
  kable(caption = "Runoff, total rainfall and rainwater harvested in each scenario")

@@

```{r, fig.height=3}

@@

map_dfr(scenarios, no2_seq, .id="Scenario") |> 
  kable(caption = "Sequestration of nitrogen dioxide in each scenario")
```

@@

scenarios |> 
  map_dfr(food_production, .id = "Scenario") |> 
  mutate(across(where(is.numeric), ~ .x/1000)) |> 
  kable(caption = "Food production in Tm/year in each scenario")

@@

## Acknowledgements {.unnumbered}

We want to aknowledge the `R` community for being so supportive and friendly in StackOverflow, but also in Twitter and Mastodon under the hashtag #Rstats. Also, to Hadley Wickham and his team, for all the tools that make the life easier to R developers. Likewise, we want to thank all R developers that altruistically spend their time in building open-source packages.
